Reproducible Research
======================

1. Loading and preprocessing the data.
----------------------
Read data from activity.csv, with column class *Integer, Date and Integer*.

```{r, echo=TRUE}
data <- read.csv("activity.csv",
                 colClass = c("integer", "Date", "integer"))
summary(data)
str(data)
```
2. What is mean total number of steps taken per day?
----------------------

First, lets add up number of steps taken each day.

```{r, echo=TRUE}
steps.each.day <- tapply(data$steps, data$date, sum, na.rm = TRUE)
```

Then, we load ggplot2 library and use qplot to draw a histgram with 30 bins.

```{r, echo=TRUE}
library(ggplot2)
hist1 <- qplot(steps.each.day, binwidth = (max(steps.each.day) - min(steps.each.day)) / 30)
print(hist1)
```

Finally, the mean and median is calculated as below.

```{r, echo=TRUE}
mean.steps <- mean(steps.each.day)
median.steps <- median(steps.each.day)
```

We got: mean.steps = `r mean.steps`, and median = `r median.steps` 

3. What is the avergae daily activity pattern?
----------------------

First we calculate mean steps each interval.

```{r, echo=TRUE}
mean.steps.per.interval <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
```
Then we draw the line graph.

```{r, echo=TRUE}
interval = as.numeric(rownames(mean.steps.per.interval))
qplot(interval, mean.steps.per.interval, geom = "line")
```

Finally, we find the maximum.

```{r, echo=TRUE}
index.max <- as.numeric(rownames(mean.steps.per.interval)[
             which(mean.steps.per.interval == max(mean.steps.per.interval))])
hour <- floor(index.max / 100)
minute <- index.max - hour * 100
```

Interval `r index.max` contain the maximum number of steps.
It is `r hour`:`r minute`

4. Inputing missing values
----------------------

First we use package **mice** to look for missings.

```{r, echo=TRUE}
library(mice)
md.pattern(data)
num.missings <- sum(is.na(data$steps))
```

There are `r num.missings` missing steps in activity.csv.

Then we fill the missings with mean of the same interval across all days.

```{r, echo=TRUE}
find.mean <- function(interval) {
  mean.steps.per.interval[as.character(interval)]
}
complete.data <- data
for (i in 1:nrow(complete.data)){
  if(is.na(complete.data$steps[i])){
    complete.data$steps[i] <- find.mean(complete.data$interval[i])
  }
}
md.pattern(complete.data)
```

Histgram of the complete data is shown below.

```{r, echo=TRUE}
steps.each.day.complete <- tapply(complete.data$steps, complete.data$date, sum, na.rm = TRUE)
hist2 <- qplot(steps.each.day.complete, binwidth = (max(steps.each.day.complete) - min(steps.each.day.complete)) / 30)
print(hist1)
print(hist2)
mean.steps.complete <- mean(steps.each.day.complete)
median.steps.complete <- median(steps.each.day.complete)
```

We got:   mean.steps = `r mean.steps.complete`, and median = `r median.steps.complete` 

Original: mean.steps = `r mean.steps`, and median = `r median.steps` 

By comparison, we found that as we replace missings with mean, daily activity is closer to a gaussian distribution. 


5. Are there differences in activity patterns between weekdays and weekends:
----------------------

Here we use package **reshape** to group Original data by daytype.

```{r, echo=TRUE}
complete.data$daytype <- format(complete.data$date, "%u")
complete.data$daytype <- complete.data$daytype > 5
library(reshape2)
week.data <- melt(complete.data, id = c("interval", "daytype", "date"),
                  measure.vars = c("steps"))
week.data <- dcast(week.data, interval + daytype ~ variable, mean) 
week.data$daytype <- factor(as.integer(week.data$daytype), label = c("Weekday", "Weekend"))
```

The final graph shown here.

```{r, echo=TRUE, fig.width=10}
qplot(interval, steps, data = week.data, geom = "line", facets = daytype~.)
```
